%!TEX root = ../thesis.tex

\chapter{First Appendix}
\label{sec:first_appendix}

\section{GitHub repositories}

All code developed in this project is published on GitHub and can be found at the following links.

\begin{itemize}
	\item{Code and documentation related to the local communication gateway running on a Raspberry Pi 2:\\ \url{https://github.com/spiegelm/smart-heating-local}}
	
	\item{Code and documentation related to the server:\\ \url{https://github.com/spiegelm/smart-heating-server}}
	
	\item{Code and documentation related to the mobile application:\\ \url{https://github.com/Octoshape/smart-heating-app}}
\end{itemize}

\section{Setup instructions for the local communication gateway}

This appendix explains the setup routine of the local communication gateways implemented on a Raspberry Pi 2.
See also the repository documentation:\\
\url{https://github.com/spiegelm/smart-heating-local}

\subsection{Setup raspbian}\label{setup-raspbian}

Flash raspbian to a SD card and boot the Raspberry. Find the IP address
using \highlight{nmap -sP [ip-adress]/[bitmask]}, e.g.
\highlight{nmap -sP 192.168.0.0/24}.

Open a SSH client and connect to the determined IP. The default username
and password are \highlight{pi} and \highlight{raspberry}. Type
\highlight{sudo raspi-config} to expand the filesystem, change the
password and set the local time zone.

Enable IPv6: Add \highlight{ipv6} on a line by itself at the end of
/etc/modules.

\subsection{Install dependencies}\label{install-dependencies}

Install \highlight{at}. Needed to remain tunslip6 started because UDEV
rules kill the spawning process.

\begin{verbatim}
sudo apt-get install at
\end{verbatim}

\subsubsection{Install Python 3.4.1 and	aiocoap}\label{install-python-3.4.1-and-aiocoap}

Credits to Marc Hüppin for the initial version.

openssl and libssl-dev are required for SSL support in python and is also required by pip.

\begin{quote}
	\highlight{sudo apt-get install sqlite3 libsqlite3-dev openssl libssl-dev}
	
	install the sqlite3 packages
	
	\begin{verbatim}
	mkdir ~/src
	cd ~/src
	wget https://www.python.org/ftp/python/3.4.1/Python-3.4.1.tgz
	\end{verbatim}
	
	unpack cd into dir
	
	\begin{verbatim}
	./configure
	make
	sudo make install
	\end{verbatim}
	
	get the aiocoap library from github: https://github.com/chrysn/aiocoap
	
	get setuptools from: https://pypi.python.org/pypi/setuptools
	
	install aiocoap using
	
	\begin{verbatim}
	python3.4 setup.py install
	\end{verbatim}
\end{quote}

\subsection{Setup smart-heating-local}\label{setup-smart-heating-local}

\subsubsection{Setup the border router connection}\label{setup-the-border-router-connection}

Clone this repository into your home folder:

\highlight{git clone https://github.com/spiegelm/smart-heating-local.git}.

Create symbolic links

\begin{itemize}
	\item{udev rules: \highlight{sudo ln -s /home/pi/smart-heating-local/rules.d/90-local.rules /etc/udev/rules.d/}}
	\item{tunslip executable:\\
		\highlight{sudo ln -s /home/pi/smart-heating-local/bin/tunslip6 /bin/}}
\end{itemize}

Add this line to \highlight{/etc/rc.local} to make sure the udev rule is also executed on startup

\noindent
\begin{minipage}{\linewidth}
	\begin{lstlisting}[numbers=none]
udevadm trigger --verbose --action=add --subsystem-match=usb --attr-match=idVendor=0403 --attr-match=idProduct=6001
	\end{lstlisting}
\end{minipage}

Reboot: \highlight{sudo reboot}

Attach the sky tmote usb dongle to the raspberry. The tun0 interface should be shown by \highlight{ifconfig}.
In case of problems run \highlight{sudo \textasciitilde/smart-heating-local/bin/\allowbreak start\_tunslip.sh}
manually.
Determine the ipv6 address of the Web service:
\highlight{less /var/log/tunslip6}:

\noindent
\begin{minipage}{\linewidth}
	\begin{lstlisting}[numbers=none]
Server IPv6 addresses:
fdfd::212:7400:115e:a9e5
fe80::212:7400:115e:a9e5
	\end{lstlisting}
\end{minipage}

Retrieve the registered routes on the border router:

\highlight{wget http://[fdfd::212:7400:115e:a9e5]}:

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}[numbers=none]
<html><head><title>ContikiRPL</title></head><body>
Neighbors<pre>fe80::221:2eff:ff00:22d3
</pre>Routes<pre>fdfd::221:2eff:ff00:22d3/128 (via fe80::221:2eff:ff00:22d3) 16711422s
</pre></body></html>
\end{lstlisting}
\end{minipage}

Test route to the thermostat by requesting the current temperature via
coap-client (libcoap):

\highlight{\textasciitilde/smart-heating-local/bin/coap-client -m get\\
coap://[fdfd::221:2eff:ff00:22d3]/sensors/temperature}

Congratulations, your raspberry is connected to a thermostat!

\subsubsection{Install the required python packages}\label{install-the-required-python-packages}

Install pip: \url{https://pip.pypa.io/en/latest/installing.html}

\noindent
Install the project requirements:

\noindent
\highlight{cd ~/smart-heating-local/}\\
\highlight{pip install -r requirements.txt}

\subsubsection{Configure cron tasks}

Setup crontab to run the log and upload scripts periodically:

%\noindent
\highlight{crontab -e}

Insert these lines at the end of the file:

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}
*/15 * * * * /usr/local/bin/python3.4 /home/pi/smart-heating-local/thermostat_sync.py
*/5 * * * * /usr/local/bin/python3.4 /home/pi/smart-heating-local/server_sync.py
\end{lstlisting}
\end{minipage}

These commands ensure that the temperature is polled from the registered
thermostats each 15 minutes and checked for uploading to the server each
5 minutes. The scripts log interesting events to
\highlight{\textasciitilde/smart-heating-local/logs/smart-heating.log}.


\section{Setup instructions for the server infrastructure}

Setup a virtual or dedicated server based on Ubuntu Linux 14.04 LTS and enter the following the commands in a terminal:

\noindent
\begin{minipage}{\linewidth}
	\begin{lstlisting}
# Install dependencies
sudo apt-get update
sudo apt-get upgrade
sudo apt-get install git python-virtualenv python3-pip
# Install code into home directory
cd ~
git clone https://github.com/spiegelm/smart-heating-server
# Setup virtualenv
virtualenv -p python3 env
. ~/env/bin/activate
# Check for python version >=3.4.0
python --version
# Install requirements, setup database and start server
cd ~/smart-heating-server/
pip install -r requirements.txt
./manage.py migrate
# Run server on port 8000
./manage.py runserver 0.0.0.0:8000
# Test that server is accessible via browser
# Kill server: CTRL-C
# Start server in the background using nohup
./scripts/restart_server.sh
# Kill background server
./scripts/kill_server.sh
	\end{lstlisting}
\end{minipage}

In case of software updates run the following commands to fetch the newest code form the repository and restart the server:

\noindent
\begin{minipage}{\linewidth}
	\begin{lstlisting}
cd ~/smart-heating-server
git pull
./scripts/restart_server.sh
	\end{lstlisting}
\end{minipage}



\section{Setup instructions for the mobile application}

To install the application on your Android device simply download the apk file from \url{https://github.com/Octoshape/smart-heating-app/blob/master/smartheating-debug.apk} and transfer it onto your device. Then you can install the application from the apk file.

The application will ask for permissions for the NFC adapter (to be used for scanning the NFC tags), permission for internet usage, for obvious reasons and to access the network state in order to find out if your device is connected to the internet.

The minimal required API is Android 4.1 (JELLY\_BEAN), the targeted API is Android 5.0 (LOLLIPOP).








