%!TEX root = ../thesis.tex

\chapter{Future Work}
\label{sec:futurework}

%Anything that is interesting -- but you did not have the time to pursue -- fits into future work.

In the scope of this lab project we developed a reliable heating control infrastructure and the corresponding innovative mobile application.
We focused these implementations to build a solid basis for future projects to build upon.
In the following we outline a few interesting ideas that we did not have the time to pursue but that could be implemented as a future work.

%The developed Web server builds a solid tested framework that stores and processes the collected data.
The developed system could be extended to track and predict user locations.
This would allow to dynamically adjust the heating schedule to react to spontaneous user behavior deviations and could further decrease the energy consumption and increase the users comfort level in certain cases.

As a part of this project we implemented a mobile application designed for Android smart phones.
Another possibility for future work would be an additional user interface that could show more complex data analysis suitable for devices with bigger screens.

For the mobile application itself we were able to come up with many ideas for new features which we did not have the time to pursue. Some of these are:

\begin{itemize} 
	\item{\textbf{Schedule copying}}
	
	The ability to copy a schedule from a different room, so the user does not have to change every room's schedule separately. Or even add the ability to change multiple schedules simultaneously.
	\item{\textbf{Personalized schedules}}
	
	Each user has its own personalized schedule, and the rooms can be assigned to certain users. If only a single user is assigned to a room, his personalized schedule is used in that room, but if multiple users are assigned to the same room (e.g. a living room) then the application calculates an average of all the different schedules or tries to find the best schedule to suit everyone's needs.
	\item{\textbf{Home view features}}
	
	The home view can be extended in many ways. For example rearranging the rooms to the user's liking, grouping of rooms to resemble the actual layout of the residence more or even being able to freely move the rooms such that each level of the residence can be laid out like the actual building. This would help the user find the rooms he is looking for much more quickly and easily.
	\item{\textbf{User profile}}
	
	Create user profiles with interesting statistics like an estimate of how much money was saved by using the application, how long the heating was turned completely off this year, and so on.
\end{itemize}